---

title: "usmap"
author: "Laura Ellis"
date: "11/27/2019"
output:
  md_document:
    variant: markdown_github
---

## Tutorial 

The full tutorial and description is available at [littlemissdata.com](https://www.littlemissdata.com/blog)

## Install and Load Packages

Check for packages installed, install if needed and load package.  I found this code off of
[Vikram Baliga's blog](https://www.vikram-baliga.com/blog/2015/7/19/a-hassle-free-way-to-verify-that-r-packages-are-installed-and-loaded)

```{r}
#specify the packages of interest
packages = c("gtrendsR","tidyverse","usmap")

#use this function to check if each package is on the local machine
#if a package is installed, it will be loaded
#if any are not, the missing package(s) will be installed and loaded
package.check <- lapply(packages, FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
        install.packages(x, dependencies = TRUE)
        library(x, character.only = TRUE)
    }
})

#verify they are loaded
search()

#Set some variables

orange <- "#C9592E"

```


## Get thanksgiving trends

Use the [gTrendsR Package](https://cran.r-project.org/web/packages/gtrendsR/gtrendsR.pdf) to get the query trends for thanksgiving in the US for the past 24 hours

```{r}
thanksgiving <- gtrends("thanksgiving",geo = "US", time = "now 1-d") # last day
```

## Gather Interest by State

Select the data frame for interest by state and then convert the names of the states to FIPS codes (2 characters for state, 5 characters for county).

```{r }

thanksgivingStates <- thanksgiving$interest_by_region

#The usmap requires the data in FIPS format.
# To gather the FIPS data, use the fips() function to convert state name to code
#https://www.rdocumentation.org/packages/cdlTools/versions/0.14/topics/fips

thanksgivingStates$fips <-fips(thanksgivingStates$location)

```

## Plot interest with the US Map by state

Create a US heatmap with google search popularity for the keyword "thanksgiving".

```{r }

plot_usmap(data = thanksgivingStates, values = "hits",  color = orange, labels=TRUE) + 
  scale_fill_continuous( low = "white", high = orange, 
                         name = "Popularity", label = scales::comma
  ) + 
  theme(legend.position = "right") + 
  theme(panel.background = element_rect(colour = "black")) + 
  labs(title = "Popularity of Thanksgiving Google Search by State", caption = "Source: @littlemissdata") 
  
```



## Plot interest with the US Map by state

Drill in on the seemingly most popular regions using the "include" parameter in the plot_usmap() function.  Regional divisions can be found in the docs [here](https://cran.r-project.org/web/packages/usmap/usmap.pdf)

```{r }

plot_usmap(data = thanksgivingStates, values = "hits", include =  c(.south_atlantic, .east_north_central), color = orange, labels=TRUE) + 
  scale_fill_continuous( low = "white", high = orange, 
                         name = "Popularity", label = scales::comma
  ) + 
  theme(legend.position = "right") + 
  theme(panel.background = element_rect(colour = "black")) + 
  labs(title = "US East Coast Popularity of Thanksgiving Google Search", caption = "Source: @littlemissdata") 
  
```


