---
title: "Patchwork"
author: "Laura Ellis"
date: "3/2/2020"
output:
  md_document:
    variant: markdown_github
---

# Set Up

## Install and load all the packages

Thank you to "Dusty" who posted the package "easypackages" on my [last tutorial](https://www.littlemissdata.com/blog/bacheloranalysis)

```{r}
#install.packages("easypackages")
library(easypackages)
packages("tidyverse", "rtweet", "tidytext", "rtweet", "wordcloud2", "patchwork", "cran.stats", "data.table", 
         "gameofthrones", "ggimage", "magick", "ggpubr", "jpeg", "png")

```


## Set up our colour palette

Using the beautiful Game of Thrones color palette from [Alejandro Jiménez](https://twitter.com/aljrico) in his ["gameofthrones" package.](https://github.com/aljrico/gameofthrones).  Thank you to [Divya Seernani](https://twitter.com/DSeernani) for sharing!
```{r}

#Set the palette using the beautiful GOT Arya palette from Alejandro Jiménez
pal <- got(20, option = "Arya")

#cherry pick a few extended
c <-"#889999"
c2 <- "#AAB7AF"
```

## Add your twitter credentials
Create your twitter authentication token by following the steps in [Michael Kearneys](https://mikewk.com/) [beautiful documentation](https://rtweet.info/articles/auth.html).  Replace the "ADD YOUR CREDS" with your own credentials. 

```{r}
#create_token(
#  app = "ADD YOUR CREDS",
#  consumer_key = "ADD YOUR CREDS",
#  consumer_secret = "ADD YOUR CREDS")
```

# Create Plots

## 1st Plot - Create a plot of the tweet stats (favorites, retweets)

### Lookup the tweet and view stats


```{r}
lt <-lookup_tweets('1229176433123168256')
lt
```

### Create a chart with the tweet stats

```{r}
p1 <- lt %>% 
  rename(Faves = favorite_count, RTs = retweet_count) %>% 
  select(Faves, RTs) %>%  #select only the desired columns
  gather("stat", "value") %>%  #reformat to make the table long which is easier for bar charts to consume
  ggplot(aes(stat, value)) +  #plot the bar chart
  geom_bar(stat="identity", fill=c2) +
  theme_classic() + 
  labs(title = "Tweet Stats",
                           x = "Tweet Statistic", y = "Total")  

p1
```


## 2nd Plot - Create a plot of the tweet stats (favorites, retweets)

### Gather approx 1K of the retweet data

The [get_retweets()](https://rtweet.info/reference/get_retweets.html) function only allows a max of 100 retweets to be pulled via the API at a time. This is a rate imposed by the twitter API.  When pulling this data, I had quite a difficult time.  Not only, did a lot of the suggested methods to getting cursors fail, the rate limiting wasn't consistent.  Sometimes I was able to get close to 1K tweets in 100 batches.  Sometimes it blocked me for 15 min intervals (as expected).  Since this is just an example to show patchwork, I decided to just grab 1K of the retweets which is roughly half of the full set of retweets.  Further, I should let you know that I did attempt to put it in a function, but I couldn't find an appropriate system wait time that would complete in a reasonable time and/or actually return the data.  Please reach out if you have a better/proven method! In the meantime, here is my brute force method.

```{r}

statusid <- '1229176433123168256' #set the first lowest retweet statusid to be the id of the original tweet
rtweets <- get_retweets(statusid, n=100, parse = TRUE) # get 100 retweets
min_id <- min(rtweets$status_id)

rtweets2 <- get_retweets(statusid, n=100, max_id = min_id, parse = TRUE) # get 100 retweets
min_id <- min(rtweets2$status_id)

rtweets3 <- get_retweets(statusid, n=100, max_id = min_id, parse = TRUE) # get 100 retweets
min_id <- min(rtweets3$status_id)

rtweets4 <- get_retweets(statusid, n=100, max_id = min_id, parse = TRUE) # get 100 retweets
min_id <- min(rtweets4$status_id)

rtweets5 <- get_retweets(statusid, n=100, max_id = min_id, parse = TRUE) # get 100 retweets
min_id <- min(rtweets5$status_id)

rtweets6 <- get_retweets(statusid, n=100, max_id = min_id, parse = TRUE) # get 100 retweets
min_id <- min(rtweets6$status_id)

rtweets7 <- get_retweets(statusid, n=100, max_id = min_id, parse = TRUE) # get 100 retweets
min_id <- min(rtweets7$status_id)

rtweets8 <- get_retweets(statusid, n=100, max_id = min_id, parse = TRUE) # get 100 retweets
min_id <- min(rtweets8$status_id)

rtweets9 <- get_retweets(statusid, n=100, max_id = min_id, parse = TRUE) # get 100 retweets
min_id <- min(rtweets9$status_id)

rtweets10 <- get_retweets(statusid, n=100, max_id = min_id, parse = TRUE) # get 100 retweets
min_id <- min(rtweets10$status_id)

rtweet_table <-rbind(rtweets, rtweets2, rtweets3, rtweets4, rtweets5, rtweets6, rtweets7, rtweets8, rtweets9, rtweets10)
nrow(rtweet_table)
length(unique(rtweet_table$description))

```

### Graph the most common words used in the retweeters profile descriptions

Most of the techniques used below to process the data and graph the data are taken from the [tidy text mining](https://www.tidytextmining.com/) book by [Julia Silge](https://twitter.com/juliasilge) and [David Robinson](https://twitter.com/drob)


```{r}

data(stop_words)
#Unnest the words - code via Tidy Text
rtweet_table2 <- rtweet_table %>% 
  unnest_tokens(word, description) %>% 
  anti_join(stop_words) %>% 
  count(word, sort = TRUE) %>%
  filter(!word %in% c('t.co', 'https'))

wc <-wordcloud2(rtweet_table2, size=0.7, color=rep_len(pal, nrow(rtweet_table2) ) )
wc 

p2 <- rtweet_table2 %>%
  filter(n> 50) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n)) +
  theme_classic() +
  geom_col(fill= c) +
  labs(title = "RT Profiles",
       x = "Key Words", y = "Total Occurances")  +
  coord_flip() 

p2
```



## 3rd Plot - Plot the patchwork CRAN download stats

### Gather the data

To download the patchwork download stats, I used the ["cran.stats" package](https://www.rdocumentation.org/packages/cran.stats/versions/0.1).  The examples to process the download stats were very easy and I just followed the [examples here](https://cran.r-project.org/web/packages/dlstats/vignettes/dlstats.html#download-stats-of-cran-packages).

```{r}

dt = read_logs(start = as.Date("2020-02-01"), 
               end = as.Date("2020-02-29"), 
               verbose = TRUE)
dim(dt)

patchwork <- stats_logs(dt, type="daily", packages=c("patchwork"), 
                        dependency=TRUE, duration = 30L)
```


### Plot the CRAN download data

Using the ggplot, geom_line() function with just a little extra fancyness to annotate the graph with the annotate() function.  Great annotation examples [here](https://ggplot2.tidyverse.org/reference/annotate.html)

```{r}
p3 <- ggplot(patchwork, aes(x=key, y=tot_N, group=1)) +
  geom_line() + theme_classic() + theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  ylim(0, 1500) + 
  labs(title = "Downloads of the R Patchwork Package",
       x = "Date", y = "Total Downloads") + 
  annotate("rect", xmin = "2020-02-16", xmax = "2020-02-20", ymin = 400, ymax = 900,
           alpha = .3, fill = c2)  +
  annotate(
    geom = "curve", alpha = 0.3, x = "2020-02-14", y = 650, xend = "2020-02-17", yend = 800, 
    curvature = .3, arrow = arrow(length = unit(2, "mm"))
  ) +
  annotate(geom = "text", x = "2020-02-07", y = 650, label = "Nerd viral #rstats tweet", hjust = "left", alpha = 0.5)

p3

```


# Add the plots to the same graphic using patchwork

As is the focus of this post, when this package was [shared on twitter](https://twitter.com/LittleMissData/status/1229176433123168256), people were very excited about it.  The [patchwork package](https://github.com/thomasp85/patchwork) was created by [Thomas Lin Pedersen](https://twitter.com/thomasp85).  Not only is it incredibly easy to use, it comes with great [documentation](https://patchwork.data-imaginist.com/index.html)

## Try a few layouts

Using the plots p1, p2, p3 created above, try a few layouts following the package documentation

```{r}
p1/p2/p3
p1 + p2 + p3
p1/ (p2 +p3)

#Final Layout
p <- p3 / (p1 + p2)
p
```

## Annotate the final layout

We will select the final layout from the above code block and then add some overall titles, captioning and formatting.  This example was covered in the excellent [patchwork annotation guide](https://patchwork.data-imaginist.com/articles/guides/annotation.html)

```{r }
# Change the values for the labels
# Overall annotation

p + plot_annotation(
  title = 'Patchwork Went Nerd Viral',
  caption = 'Source: @littlemissdata'
) 


p + plot_annotation(
  title = 'Patchwork Went Nerd Viral',
  caption = 'Source: @littlemissdata'
) & 
  theme(text = element_text('mono'))
```

# Add an image to the patchwork graphic

## Bring in the image

Using an empty plot and the background_image() function, you can bring an image into a graph object.  Further, you can prevent image resizing with the coord_fixed() function.  This is important so the actual image doesn't get resized with the patchwork placement. 

```{r}
## Add an image

twitter <- image_read('https://raw.githubusercontent.com/lgellis/MiscTutorial/master/Patchwork/twitter_post.png')
twitter <- ggplot() +
  background_image(twitter) + coord_fixed()
```

## Plot the image with patchwork

```{r }

(twitter + p3) / (p1 + p2 )
pF <- twitter + (p3/ (p1 + p2))

pF + plot_annotation(
  title = 'Patchwork Went Nerd Viral',
  caption = 'Source: @littlemissdata'
) & 
  theme(text = element_text(family ='mono', size = 9))

```
































