---
title: "Happy Holidays 2019"
author: Laura Ellis
date: Dec 24 2019
output:
  md_document:
    variant: markdown_github
---

# Load all packages

```{r}
library(data.table)
library(tidyverse)
library(gganimate)
library(ggimage)
library(magick)
library(gifski)
```

# Download the data

Csv created with a tool called [coordinator](https://spotify.github.io/coordinator/), by [Aliza Aufrichtig](https://twitter.com/alizauf).  It allows us to convert SVG graphics to data co-ordinates

```{r}
df= fread('https://raw.githubusercontent.com/lgellis/MiscTutorial/master/Holidays/Holidays.csv', stringsAsFactors = FALSE)
```

#Create the Plot

Classic ggplot scatterplot.  Remove all lines an chart aspects with theme_void.  Increase the chart size beyond the plotted points using the xlim() and ylim() functions.  This will allow us to nicely center the text on our background image. 

```{r}
# Basic scatter plot
p <-ggplot(df, aes(x=x, y=y)) +
  geom_point(colour = '#BF0000', size = 0.3) +
   xlim(0, 18000) +
   ylim(0, 10000) +
   theme_void()
p

```


# Add the background Holiday Image

Add the holiday background using the ggbackground funcion in the ggimage package.  I got this tip from a [great blog post](https://guangchuangyu.github.io/2018/04/setting-ggplot2-background-with-ggbackground/) by [Guangchuang Yu](https://guangchuangyu.github.io/)

```{r}

img <- "https://raw.githubusercontent.com/lgellis/MiscTutorial/master/Holidays/%20background.jpg"
imgRead <- image_read(img)
ggbackground(p, img)

```



# Animation

Create the animation with the gganimate package.  Use transition_manual vs transition_reveal because it allows us to specify cumulative=TRUE to keep all data already shown.  


```{r}
t <-ggplot(df, aes(x=x, y=y)) +
   theme_void() + 
   theme(plot.background = element_rect(fill = '#CA302F')) +
   geom_point(colour = 'white', size = 1) +
   transition_manual(as.numeric(x), cumulative=TRUE) 

```


# Save a gif

Use the gifski package to save the gif. 

```{r}

gif <- animate(t, end_pause = 25, width = 800, height = 400, fps = 8)
gif
anim_save("HappyHolidays.gif", gif)

```


