---
title: "EDA Notebook"
output:
  word_document: default
  html_notebook: default
---

This is a notebook used to explore data. 

#Use Data Explorer to Get a Greater Summary View

```{r}
#Bring in the data set

library(readr)

df= read_csv('https://raw.githubusercontent.com/lgellis/STEM/master/DATA-ART-1/Data/FinalData.csv', col_names = TRUE)

```


#Use the head command to get a preview of the data set

The default is 6 rows, but let's select 10 just to get a bigger picture

```{r}
head(df, 10)
```

```{r}
dim(df)

#Displays the type and a preview of all columns as a row so that it's very easy to take in.

library(dplyr)
glimpse(df)


```


#Summary Statistics

This is a good method because it shows the summary statistics for numerical values

```{r}
summary(df)
```


#Use Skimr to Display Summary Statistics for the variables

This has same as above + missing and a histogram.  Also, it has some additional statistics of non-numerical values.
```{r}
library(skimr)
skim(df)
```


# Get a Full Data Summary Report

Basic statistics, missing data, distribution, correlation, PCA.

```{r}

library(DataExplorer)
DataExplorer::create_report(df)

```










