---
title: "Bachelor Ratings"
author: Laura Ellis
date: Sept 24, 2019
output:
  md_document:
    variant: markdown_github
---

#Introduction

In this blog post, I will be analyzing the relative popularity of the bachelor franchise seasons as measured by their google search popularity.  

In this tutorial I will be using the [gtrendsR package](https://cran.r-project.org/web/packages/gtrendsR/gtrendsR.pdf) to gather Google trend information, dplyr to format the data, ggplot2 to create the graphs, gganimate to make an animation and geom_image to create custom lollipop charts.  


#Load all of the packages and install if necessary

In my [latest blog post](https://www.littlemissdata.com/blog/iconmap), someone kindly suggested that I do an auto check to install all necessary packages before loading them.  After a quick search, I found this code below to efficiently install and load packages in [Vikram Baliga's Blog](https://www.vikram-baliga.com/blog/2015/7/19/a-hassle-free-way-to-verify-that-r-packages-are-installed-and-loaded)

```{r}
#specify the packages of interest
packages = c("gtrendsR","tidyverse","gifski", "gganimate", "ggimage", "lubridate")

#use this function to check if each package is on the local machine
#if a package is installed, it will be loaded
#if any are not, the missing package(s) will be installed and loaded
package.check <- lapply(packages, FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
        install.packages(x, dependencies = TRUE)
        library(x, character.only = TRUE)
    }
})

#verify they are loaded
search()
```
 
# Set the color variables
 
To ensure consistent and effective color formatting, I am setting the color variables up front. 

```{r}
pink <- "#FF8DC6"
blue <- "#56C1FF"
yellow <- "#FAE232"
```
 
# Gather the ratings data for: The Bachelor, The Bachelorette and Bachelor in Paradise

Using the gtrendsR package, load weekly US ratings for "Bachelor in Paradise", "The Bachelor" and "The Bachelorette".  The hits are calculated with a relative max of 100 to show the relative max hits over the time range and search subjects.  

Plot the trends with the plot() function.

```{r}
bachTrends <- gtrends(c("Bachelor in Paradise", "The Bachelor", "The Bachelorette"), geo ="US")
plot(bachTrends)
bachTrendsInterest <- bachTrends$interest_over_time
```

# Transform the data

Filter to data that is 2017+
Convert hits to numeric as the default is character


```{r}
trends <- bachTrendsInterest %>% 
  filter(year(date)>2016) %>% 
  mutate(date = ymd(date),
         hits = as.numeric(hits))
```

# Create the same plot with ggplot2

Create the basic plot of relative search popularity by search critera with the ggplot2 package.  Transitioning to ggplot2 in order to use the ggplot2 features and complimentary packages like ggimage and gganimate.

```{r}
#Frequency plot by keyword
p <- ggplot() + 
  geom_line(data=trends, aes(x=date, y=hits, group=keyword, color = keyword)) + 
  scale_color_manual(values=c( yellow, blue, pink)) +
  theme_classic() +
  theme(legend.position="bottom") +
  labs(title = "The Bachelor Franchise Popularity ",
       subtitle = "Using data to find the most dramatic season ever!",
       caption = "Source: @littlemissdata", 
       x = "Date", y = "Hits") 
p
```

# Create an animation

Take the basic plot and make an animation out of it with the gganimate() package. 

```{r}
t <- p + 
  transition_reveal(as.numeric(date)) 
gif <- animate(t, end_pause = 25, width = 800, height = 400, fps = 8)
gif
anim_save("Bachelor trends", gif)

```

# Bring in meta data about bachelor franchise shows

We are going to bring in a data set which has the start dates for every single season of the Bachelor franchise. We will then do some data munging to find the closest ratings date to the season start date.  With this info we will join the bachelor season metadata to the ratings table. 

```{r}
## Add lollipops
x <-read.csv("https://raw.githubusercontent.com/lgellis/MiscTutorial/master/gtrendsR/bachelorListing.csv", stringsAsFactors = FALSE)

# Turn the dates into proper dates.  
#Ratings are only tracked on sundays so get the closest Sunday for ratings
x <-x %>% 
  mutate(startDate = ymd(as.Date(startDate, "%m/%d/%y")),
         endDate = ymd(as.Date(endDate, "%m/%d/%y")),
         ratingStartDate = floor_date(startDate, "weeks"), 
         ratingEndDate = floor_date(endDate, "weeks"))
x

#Ratings are typically highest at the beginning
x<-left_join(x, trends, by = c("topic"= "keyword", "ratingEndDate"="date"))

```

# Get the images for each of the seasons

I have some plans to layer on a lollipop graph with the image of the bachelor season and display it at the height of the number of hits on the last day of the season.  As such, I need to assign an image to every single season.  

```{r}

x <-x %>% 
  mutate(Image = case_when(season == "Nick Viall" ~ "https://raw.githubusercontent.com/lgellis/MiscTutorial/master/gtrendsR/images/Nick.png",
                           season == "Arie Luyendyk Jr" ~ "https://raw.githubusercontent.com/lgellis/MiscTutorial/master/gtrendsR/images/Arie.png", 
                           season == "Colton Underwood" ~ "https://raw.githubusercontent.com/lgellis/MiscTutorial/master/gtrendsR/images/Colton.png", 
                           season == "Rachel Lindsay" ~ "https://raw.githubusercontent.com/lgellis/MiscTutorial/master/gtrendsR/images/Rachel.png", 
                           season == "Becca Kufrin" ~ "https://raw.githubusercontent.com/lgellis/MiscTutorial/master/gtrendsR/images/Becca.png", 
                           season == "Hannah Brown" ~ "https://raw.githubusercontent.com/lgellis/MiscTutorial/master/gtrendsR/images/Hannah.png", 
                           topic == "Bachelor in Paradise" ~ "https://raw.githubusercontent.com/lgellis/MiscTutorial/master/gtrendsR/images/BIP.png"))


```


# Create the mega chart

Create a fun graph to display the relative ratings for each season by layering on a lollipop chart to represent the seasons and their relative search popularity for the last week of the season.  Use the geom_segment() function to set the lollipop stem and the geom_image() function to set the lolipop circle with the image representing the season.  

```{r, fig.width = 8, fig.height=8}
p <- ggplot() + 
  geom_line(data=trends, aes(x=date, y=hits, group=keyword, color = keyword), size=1) + 
  scale_color_manual(values=c(yellow, blue, pink)) +
  geom_segment(data=x, aes(x=ratingEndDate, 
                           xend=ratingEndDate, 
                           y=0, 
                           yend=hits, 
                           color=topic), size=1) +
  geom_image(data=x, aes(x=ratingEndDate, y=hits, image=Image), size=0.105) +
  theme_classic() +
  labs(title = "The Bachelor Franchise Popularity ",
       subtitle = "Using data to find the most dramatic season ever!",
       caption = "Source: @littlemissdata", 
       x = "Date", y = "Hits") +
  theme(legend.position="none",
    plot.title = element_text(size = 12, face = "bold"),
    plot.subtitle = element_text(size=10, face = "italic"),
    plot.caption = element_text(size = 8, face = "italic") )

p
```


